\LARGE \bf Reactive Trajectory Planning and Trajectory Tracking for Pedestrian-Aware Autonomous Driving in Urban Environments

### 2016-01-23 root.tex
final submit version from me

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
In this paper, we address the problem of trajectory planning for urban roads while reacting to the presence of pedestrians.
Past work reduces jerk and limits velocities and acceleration to smooth trajectories, but it does not consider reactive behaviors such as responding to pedestrians.
On the other hand, systems for collision avoidance plan paths around obstacles and pedestrians, stop for them, or produce a warning. 
In this paper, we present an integrated trajectory generation and tracking system for urban environments that simultaneously considers constraints to maximize human comfort and real-time trajectory updates to react to pedestrians on the road.

@@

When planning trajectories for an autonomous ground vehicle operating on roadways, a typical goal is to minimize navigation time.
% "path-trajectory" is improper. "path-velocity" is used extensively in the literature.
It is common to employ the path-velocity decomposition to separate the two tasks so that they can be carried out sequentially or in parallel.
% (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
First is {\it Path Planning}, which involves computation of the optimal path.
The second task is {\it Trajectory Planning}, which involves the computation of the fastest velocity profile that is appropriate for the scenario.

@@

For human passengers, it is widely believed that high levels of jerk (time derivative of acceleration) and high acceleration/velocity values are prime contributors to ride discomfort.
A lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
Similar work has been performed by Bianco et al \cite{GuarinoLoBianco2004,GuarinoLoBianco2005,Bianco2007,Bianco2009,GuarinoLoBianco2013}, which sets the travel time a priori and optimally minimizes jerk but does not set hard limitations or control jerk directly.
Some preliminary work on optimizing these constraints on urban roads has also been done by ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene.
On the other hand, many systems in the literature find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} but do not work as part of an integrated system that can navigate urban roads.
This approach also makes the assumption that alternative paths around pedestrians are available, which may be invalid in urban scenarios.
Planners which optimize human comfort by modifying the whole trajectory to seek obstacle-free space \cite{Villagra2012,Villagra2012a} do not feature stopping as a well-integrated behavior.
In this paper, we describe a trajectory generation system that satisfies these constraints for human comfort and is reactive to presence of pedestrians on the road. We enforce hard upper limits on jerk without requiring that travel time be known before planning occurs.
%Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done at very low speeds (5 miles/hour).
%There has also been some work in simulating vehicles and modeling uncertainty, but pedestrians are less predictable than cars 
%and can appear suddenly on the road (when someone is jaywalking for example).

We present an autonomy framework for pedestrian-aware ground vehicles.
Our objective is to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
%Desired behavior from the car includes following traffic laws at the intersections such as stop signs.
At stop signs the car should stop and wait for an appropriate amount of time.
It should also react to pedestrians - at intersections as well as on the road.
Pedestrians on the sidewalks should be ignored by the autonomous car.
Further, the pedestrians can be static or moving.
The car should stop, slow down and go around pedestrians when appropriate.

Given a desired path, we present a novel method for planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path without requiring prior knowledge of travel time.
Our method first segments the path using an arbitrary set of kinematic or dynamic constraints which may be functions of the desired path.
Each segment is then planned sequentially by choosing a piecewise constant jerk profile.
The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous and non-infinite throughout the entire path.
The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is called trajectory tracking.
% We are not doing control or tracking control. We're doing tracking.
We present an online method of revising the trajectory plan to safely avoid pedestrians that enter the roadway by slowing down or stopping when no alternative paths are available.

@@

Our system architecture is shown in Figure \ref{fig:addreact}.
There are 4 primary input components: Navigation Filter, Pedestrian Detector, a Path Planner, and a Map which provides the location of stop signs.
% nav filter
For navigation filtering we use an ADMA-G  commercial automotive GPS/INS system.
It receives RTK corrections via cell modem.
It outputs global position, global heading, body frame velocity and body frame acceleration to centimeter level accuracy.

@@

We assume that we have access to a path planner.
For example, this can be a service similar to Google Maps except that instead of road level route, it provides lane level route. 
We assume that the path planner operates independently from the trajectory planner and tracker.
We also assume that we have a high resolution digital map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
The destination is manually input into the system by the passengers.
Using the path planner, we obtain a path along lane centers from the vehicle's current location (computed by the navigation filter) to the destination.
This plan is then output as a set of global waypoints.

@@

Our High Resolution Map database is used to search for stop signs and provide their locations.
The plan waypoints corresponding to stop signs are then flagged for later processing.
We also have an In-Out Algorithm implemented via polygon intersection that can be used to ignore “safe” pedestrians
Pedestrians that are on sidewalks, off-path crosswalks, and behind planned stops as shown in Figure \ref{fig:intersect} can be considered ``safe'' and are ignored.
This is done by searching the same map database used for path planning.
Remaining pedestrians pose collision risk and the vehicle needs to reactively slow down or stop for them.
This is accomplished by sending stop message with pedestrian locations to the path manager.

Beyond these input subsystems, our system has three major components: a High-Level Planner, a Low-Level Planner, and a Software Controller.

@@

The Path Manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
Each of the states is described below.

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
  \caption{State Transition Diagram for ?????????? more detail????? }
  \label{fig:st}
\end{figure}

@@

Prescribed initial and end conditions for position, speed, and acceleration must be honored.
All planning is longitudinal; that is, speed profile generation is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
This means that the desired trajectory plan may be mapped onto one dimension using curvilinear coordinates to significantly simplify the solution.
We show how two dimensional constraints (specifically cornering speeds for this work) are expressed in one dimension to guide planning.

In the absence of pedestrians, the path manager cycles between NORMAL and PSTOP states.
In NORMAL state, the path manager send the first subpath to the trajectory planner, which then plans and executes it before coming to a stop at the end of the subpath.
In PSTOP mode, the path manager waits for a pre-set time duration before returning to NORMAL mode, where the next subpath is executed.
When pedestrians are present, the transition from PSTOP to NORMAL is delayed until there are no pedestrians detected in the range denoted by $\Delta s_{resume}$.

@@

All detected pedestrian locations are compared against the high-resolution digital map, so that pedestrians that are not on the roadway or near an upcoming crosswalk are not considered.
Then each pedestrian location is projected onto the total path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
Pedestrians that are on a different subpath than the vehicle, and behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
All the remaining pedestrians pose potential collision risks, however only the closest pedestrian is considered for RSTOP.

The distinction between momentarily braking for a pedestrian that quickly leaves the roadway and coming to a full stop needs no formal logic.
They follow the same state transition, and since all trajectory plans are extremely smooth, riders are rarely aware that any change took place.
It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure that no tracking errors occurred in the pedestrian detection algorithm.

\begin{figure}[thpb]
\centering
  \subfigure[In NORMAL mode, RSTOP is now triggered.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}}
  \\
  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Replan.png}}
  \\
  \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Resume.png}}
  \caption{Graphical representation of static and dynamic thresholds used to trigger state transitions.}
  \label{fig:react}
\end{figure}

When a pedestrian is detected, the nominal distance needed to stop, $\Delta s_{RSTOP}$, must be computed given the vehicle's current location, speed, and forward acceleration.
The procedure for doing this is the same as the procedure for planning the RSTOP trajectory later with a few steps omitted, since the planning algorithm is extremely efficient.

@@

This procedure is repeated when the nearest pedestrian location changes by a value more than some static buffer $\Delta s_{replan} < \Delta s_{buf}$ compared to the location for which the current RSTOP planned.
$\Delta s_{replan}$ value is typically set to 1 meter.
Normal operation is resumed when the distance between the vehicle and the closest detected pedestrian satisfies $s_{ped} - s_{veh} > \Delta s_{resume} + \Delta s_{RSTOP}$. The static resume buffer should be set such that $\Delta s_{resume} > \Delta s_{buf} + \Delta s_{replan}$.
The preceding quantities are depicted graphically in Fig. \ref{fig:react}.

@@

The task of Trajectory Planning is to take a 1D set of waypoints from the Path Manager, and compute a series of constant jerk intervals for motion along the length of the path.
The resulting trajectory must be time optimal and satisfy all the constraints on speed, acceleration, and jerk.
Elevation changes will not be addressed here.
Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, speed which is piecewise quadratic over time, and position which is piecewise cubic over time.

@@

In this section we describe our approach to fitting appropriate jerk profiles to complex sub-paths. 
We first break each sub-path into segments which have very simple shape.
Once this is done, each segment can be compared to a library of possible jerk profiles to find a good match.
Each segment will be defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ must be defined for each segment.
A sub-path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k_s}$, $\mathbf{a}_{m,k_s}$, and $\mathbf{j}_{m,k_s}$ where $k_S = 1 ... N_s$ and can be solved individually in sequence.

The segmentation process proceeds as follows:
\begin{enumerate} \label{asdf}
  \item \emph{Ceiling calculation}, where $v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
%$initial/final speeds, speed limit, peak acceleration and jerk 
are found for every point in the given sub-path.
  \item \emph{Clustering}, in which sub-path points are grouped sequentially and the segment boundaries are determined..
  \item \emph{Consolidation}, in which a single $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is selected for each segment.
\end{enumerate}
After this is done, the process described in next section may be applied to each of the segments so that a plan of jerk vs. time may be generated.

The ceiling calculation is a minimax problem.
Each of the 9 scalars in $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is a minimum of several maxima computed from an arbitrary set of constraints.
For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the sub-path.

% How we actually did this.
For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation \textbf{ RG: using centrifugal  force relationship ????} $v_{m,k_p}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{k_p}}$ for $k_p = 1, ..., N_p$, where $N_p$ is the number of waypoints in the sub-path.
The value $\kappa_{k_P}$ denotes the path curvature at each point.
For each sub-path point, the value $v_{m,k_p}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.

This information is used to choose the segment boundaries in the clustering step.
Any number of clustering algorithms may be used here, and prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
For the present example, we will choose a simple univariate heuristic which essentially identifies curves in the roadway using only the speed ceiling.
Each time that $v_{m,k} < SL$ becomes true, a curve is beginning, and a segment boundary is drawn.
Each time that $v_{m,k} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.

Figure \ref{fig:1Dto2DandSegmentation} shows the result of this algorithm. 
We start with a sub-path which is similar in structure to the sub-path between the stop signs in Fig. \ref{fig:subpathdivision}. This sub-path is broken up into five segments as shown in Fig. \ref{fig:1Dto2DandSegmentationa}, three shown in blue corresponding to straight line part of the sub-path and two shown in red corresponding to turns in the sub-path. As shown in Fig \ref{fig:1Dto2DandSegmentationb} we use the speed limit of 11.1 m/sec for the straight line segments and 3.0 m/sec for the turn segments to determine the segment boundaries.

@@

% need better notation here.
In the consolidation step, a single set of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
%initial/final speeds, speed limit, peak acceleration and jerk 
must be calculated for each segment using the values corresponding to each of the waypoints which comprise the segment in question.
The minimum of scalar across all constituent points is chosen.
While not time optimal, it does ensure that no constraints are violated while still keeping the number of total segments low.
In the present example, this means that in the straight portions of the subpath $v_m = SL$ and in the curves $v_m$ is set to the value required to obtain $a_{B,y}^{max}$ for the entirety of the curve.
As a result, many of the curves will have a constant speed.
For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are initially all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.

A jerk vs. time profile is fit to each segment using the algorithm in the next section.
Speed and acceleration must be continuous between each segment, but only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
The values of $a_i$ and $v_i$ for the first segment are set to the value currently reported from sensor data.
Since the final acceleration is to be set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments. 
Primary goal of this process is the continuity of position, speed, and acceleration.

@@

To introduce the process of computing a jerk profile for each segment, let us consider the case in which a vehicle needs to travel some arbitrary distance ahead.
The vehicle is moving at some small initial speed and needs to accelerate to some traveling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
Figure \ref{fig:full7phasespec} shows this trajectory over time.

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/Full7PhaseSpecVertical.png}
  \caption{
    7-phase constant jerk profile along curvilinear distance axis.
    All values needed to completely parameterize a single path segment for trajectory are labeled here.}
  \label{fig:full7phasespec}
\end{figure}

@@

We propose to define 4 other basic profiles to cover all feasible values in $\mathbf{b}$, and our algorithm judiciously chooses the appropriate profile. These 4 cases cover cases in which the 7-phase solution fails. For instance, if $L$ is too short or $v_m$ is too high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
The other profiles are explained below.

All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, traveling speed of zero, negative speeds in $\mathbf{b}$, etc).
It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
To find closed-form solutions for $[\Delta t_1, ..., \Delta t_M]$, given $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$, a symbolic solver was used \textbf{RG: can specify solver name???} .

For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
This results in a negative time interval for one or more phases.
In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
Specifically, jerk may be increased while peak acceleration is decreased.
This process is referred to herein as ``accel-jerk tuning''
If no acceptable values are found within the valid ranges of $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$, a new jerk profile must be chosen.

@@

% \subsection{Stitching Segment Solutions Together} \label{sec:unifiedsolution}

@@

For each segment with $M$ jerk phases, the following values are computed, which correspond to the instants at which jerk changes:
\begin{itemize}
  \item $M$ time durations $\Delta t_{ref}$. These must be added to the last time value from the previous segment to translate them into reference time.
  \item $M$ jerk values $j(t_{ref})$, which may be appended directly to the preceding jerk values
\end{itemize}
It is now possible to integrate $j(t)$ to determine an instantaneous trajectory at any moment in time.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
In this paper, we address the problem of trajectory planning for urban roads while reacting to the presence of pedestrians.
Past work reduces jerk and limits velocities and acceleration to smooth trajectories, but it does not consider reactive behaviors such as responding to pedestrians.
Other systems for collision avoidance plan paths around obstacles and pedestrians, stop for them, or produce a warning. 
In this paper, we present an integrated trajectory generation and tracking system for urban environments that simultaneously considers constraints to maximize human comfort and real-time trajectory updates to react to pedestrians on the road.

@@

When planning trajectories for an autonomous ground vehicle operating on roadways, the typical goal is to minimize navigation time.
% "path-trajectory" is improper. "path-velocity" is used extensively in the literature.
It is common to employ the path-velocity decomposition to separate the two tasks so that they can be carried out sequentially or in parallel.
% (i.e., path planning is performed prior to planning the speed that the vehicle will take along the chosen path).
First is {\it Path Planning}, which involves computation of the optimal path.
The second task is {\it Trajectory Planning}, which involves the computation of the fastest velocity profile that is appropriate for the scenario.

@@

For human passengers, it is widely believed that high levels of jerk (time derivative of acceleration) and high acceleration/velocity values are prime contributors to ride discomfort.
A lot of work has been performed on optimizing these constraints for highways and country roads ~\cite{ziegler14,bahram15,xu12,CHEB15CI}.
Similar work has been performed by Bianco et al \cite{GuarinoLoBianco2004,GuarinoLoBianco2005,Bianco2007,Bianco2009,GuarinoLoBianco2013}, which sets the travel time a priori and optimally minimizes jerk but does not set hard limitations or control jerk directly.
Some preliminary work on optimizing these constraints on urban roads has also been done by ~\cite{Rastelli14,Li15} but they do not consider reactive behavior such as responding to pedestrians or other dynamic objects in the scene.

Many systems in the literature find and plan paths around pedestrians, stop for them, or produce a warning ~\cite{pradalier05,benenson06,gu14,mogelmose15,johnson13} but do not work as part of an integrated system that can navigate urban roads.
These approaches also make the assumption that alternative paths around obstacles are available, which may be invalid in urban scenarios.
Planners which optimize human comfort by modifying the whole trajectory to seek obstacle-free space \cite{Villagra2012,Villagra2012a} do not feature stopping as a well-integrated behavior.
In this paper, we describe a trajectory generation system that satisfies the constraints for human comfort while being reactive to presence of pedestrians on the road. We enforce hard upper limits on jerk without requiring that travel time be known before planning occurs.
%Some work has been done for reacting to pedestrians such as by Singapore groups but it has been done at very low speeds (5 miles/hour).
%There has also been some work in simulating vehicles and modeling uncertainty, but pedestrians are less predictable than cars 
%and can appear suddenly on the road (when someone is jaywalking for example).

We present an autonomy framework for pedestrian-aware ground vehicles.
Our objective is to travel autonomously from start to destination while honoring traffic laws and avoiding pedestrians on the roadway.
%Desired behavior from the car includes following traffic laws at the intersections such as stop signs.
At stop sign, the autonomous car should stop and wait for an appropriate amount of time.
It should also react to pedestrians - at intersections as well as on the road.
Further, the pedestrians can be static or moving.
Pedestrians on the sidewalks should be ignored by the car.
The car should stop, slow down and go around pedestrians when appropriate.

Given a desired path, we present a novel method for planning longitudinal trajectories (i.e., the time derivatives of position in the body-forward direction) along that path without requiring prior knowledge of travel time.
Our method first segments the path using a set of kinematic or dynamic constraints which may be functions of the desired path.
Each segment is then planned sequentially by choosing a piecewise constant jerk profile.
The jerk profile is intelligently chosen from a set of pre-solved profiles such that acceleration is continuous and non-infinite throughout the entire path.
The resultant trajectory plan can then be easily sampled to provide a reference for real-time vehicle control, which is called trajectory tracking.
% We are not doing control or tracking control. We're doing tracking.
We present an online method for revising the trajectory plan to safely avoid pedestrians that enter the roadway by slowing down or stopping when no alternative paths are available.

@@

Our system architecture is shown in Figure \ref{fig:addreact}. We have three major components: a High-Level Planner, a Low-Level Planner, and a Software Controller.

Beyond these major components, our system has 4 primary input components: Navigation Filter, Pedestrian Detector, a Path Planner, and a Map which provides the location of stop signs.
% nav filter
For navigation filtering we use an ADMA-G  commercial automotive GPS/INS system.
It receives RTK corrections via cell modem.
It outputs global position, global heading, body frame velocity and body frame acceleration to centimeter level accuracy.

@@

We assume that we have access to a path planner.
For example, this can be a service similar to Google Maps except that instead of road level route, it provides lane level route. 
We assume that the path planner operates independently from the trajectory planner and tracker.
The destination is manually input into the system by the passengers.
Using the path planner, we obtain a path along lane centers from the vehicle's current location (computed by the navigation filter) to the destination.
This plan is then output as a set of global waypoints.

@@

We also assume that we have a High Resolution Digital Map with lane centers, stop signs, crosswalk geometry, and turning curves at intersections.
Our map is used to search for location of stop signs.
The plan waypoints corresponding to stop signs are then flagged for later processing.
We also have an In-Out Algorithm implemented via polygon intersection that can be used to ignore “safe” pedestrians
Pedestrians that are on sidewalks, behind planned stops, and off-path crosswalks as shown in Figure \ref{fig:intersect} can be considered {\it safe} and are ignored.
%This is done by searching the same map database used for path planning.
Remaining pedestrians pose collision risk and the vehicle needs to reactively slow down or stop for them.
%This is accomplished by sending stop message with pedestrian locations to the path manager.

@@

@@

Prescribed initial and end conditions for position, speed, and acceleration must be honored.
All planning is longitudinal; that is, speed profile generation is the primary focus, and we assume that the control subsystem will ensure that the vehicle follows the prescribed path.
This means that the desired trajectory plan may be mapped onto one dimension using curvilinear coordinates to significantly simplify the solution.
%We show how two dimensional constraints (specifically cornering speeds for this work) are expressed in one dimension to guide planning.
We express the 2D positions of the car (specifically cornering locations and speeds) in 1D $s$ coordinate system as described in Sec. \ref{sec:trajectorytracking} to guide planning.

The Path Manager is implemented as a discrete state machine, depicted in Fig. \ref{fig:st}.
Each of the states are described below.

\begin{figure}[thpb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/StateMachineSimple.png}
  \caption{State Transition Diagram for Path Manager. System transitions from NORMAL to PSTOP mode at each stop sign. In the presence of a pedestrian the system transitions to RSTOP.}
  \label{fig:st}
\end{figure}

In the absence of pedestrians, the path manager cycles between NORMAL and PSTOP states.
In NORMAL state, the path manager sends the first subpath to the trajectory planner, which then plans and executes it before coming to a stop at the end of the subpath.
In PSTOP mode, the path manager waits for a pre-set time duration before returning to NORMAL mode, where the next subpath is executed.
When pedestrians are present, the transition from PSTOP to NORMAL is delayed until there are no pedestrians detected in the range denoted by $\Delta s_{resume}$.

@@

All detected pedestrian locations are compared against the high-resolution digital map, so that pedestrians that are off roadway are not considered.
Then each pedestrian location is projected onto the 1D path in the same manner as described in Sec. \ref{sec:trajectorytracking}.
Pedestrians that are on a different subpath than the vehicle, or behind the vehicle are disregarded (i.e., $s_{ped} < s_{veh}$).
All the remaining pedestrians pose potential collision risks, however only the closest pedestrian is considered for RSTOP.

The distinction between momentarily braking for a pedestrian that quickly leaves the roadway and coming to a full stop needs no separate logic.
They follow the same state transition, and since all trajectory plans are extremely smooth, riders are rarely aware that any change took place.
It should be noted that there is a brief waiting period before resuming NORMAL mode from RSTOP to ensure that no tracking errors occurred in the pedestrian detection algorithm.

\begin{figure}[tb]
\centering
  \subfigure[In NORMAL mode, RSTOP is now triggered.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_NORMAL_time_to_stop.png}}
  \\
  \subfigure[RSTOP diagram showing bounds for revising the RSTOP trajectory.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Replan.png}}
  \\
  \subfigure[PSTOP diagram showing bounds for reverting to NORMAL mode.]{\includegraphics[width=1.0\columnwidth]{graphics/RSTOP_Resume.png}}
  \caption{Graphical representation of static and dynamic thresholds used to trigger state transitions.}
  \label{fig:react}
\end{figure}

When a pedestrian is detected, the nominal distance needed to stop, $\Delta s_{RSTOP}$, must be computed given the vehicle's current location, speed, and forward acceleration.
The procedure for doing this is the same as the procedure for planning the RSTOP trajectory with fewer steps, since the planning algorithm is extremely efficient.

@@

This procedure is repeated when the nearest pedestrian location changes by a value more than some static buffer $\Delta s_{replan} < \Delta s_{buf}$ compared to the location for which the current RSTOP planned.
$\Delta s_{replan}$ value is typically set to 1 meter.
Normal operation is resumed when the distance between the vehicle and the closest detected pedestrian satisfies $s_{ped} - s_{veh} > \Delta s_{resume} + \Delta s_{RSTOP}$. The static resume buffer should be set such that $\Delta s_{resume} > \Delta s_{buf} + \Delta s_{replan}$.
These quantities are depicted graphically in Fig. \ref{fig:react}.

@@

The task of Trajectory Planning is to take a 1D set of waypoints from the Path Manager, and compute a series of constant jerk intervals for motion along the length of the path.
The resulting trajectory must be time optimal and satisfy all the constraints on speed, acceleration, and jerk.
%Elevation changes will not be addressed here.
Integrating piecewise constant jerk over time yields acceleration which is piecewise linear over time, speed which is piecewise quadratic over time, and position which is piecewise cubic over time.

@@

In this section we describe our approach to fitting appropriate jerk profiles to complex sub-paths. 
We first break each sub-path into segments with simple shape.
Once this is done, each segment can be compared to a library of possible jerk profiles to find a good match.
Each segment is defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ must be defined for each segment.
A sub-path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_{k_s}$, $\mathbf{a}_{m,k_s}$, and $\mathbf{j}_{m,k_s}$ where $k_S = 1, ..., N_s$ and can be solved individually in sequence.

The segmentation process proceeds as follows:
\begin{enumerate} \label{asdf}
  \item \emph{Ceiling Calculation}, where $v_m$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
%$initial/final speeds, speed limit, peak acceleration and jerk 
are found for every point in the given sub-path.
  \item \emph{Clustering}, in which sub-path points are grouped sequentially and the segment boundaries are determined..
  \item \emph{Consolidation}, in which a single $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is selected for each segment.
\end{enumerate}
After these steps are completed, the process for fitting jerk profiles described in next section may be applied to each of the segments so that a plan of jerk vs. time may be generated.

The {\it Ceiling Calculation} is a minimax problem.
Each of the 9 scalars in $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ is a minimum of several maxima computed from an arbitrary set of constraints.
For instance, acceleration constraints may include a set of limits derived from friction coefficients at every point along the sub-path.

% How we actually did this.
For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation 
%\textbf{ RG: using centrifugal  force relationship ????} 
$v_{m,k_p}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_{k_p}}$ for $k_p = 1, ..., N_p$, where $N_p$ is the number of waypoints in the sub-path.
The value $\kappa_{k_P}$ denotes the path curvature at each point.
For each sub-path point, the value $v_{m,k_p}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.

This information is used to choose the segment boundaries in the {\it Clustering} step.
Any number of clustering algorithms may be used here, and a prudent choice of segment boundaries is arguably one of the most critical pieces of trajectory planning process.
We choose a simple univariate heuristic which essentially identifies curves in the roadway using only the speed ceiling.
Each time that $v_{m,k} < SL$ becomes true, a curve is beginning, and a segment boundary is drawn.
Each time that $v_{m,k} = SL$ is once again true, the current curve has ended, and another segment boundary is drawn.

Figure \ref{fig:1Dto2DandSegmentation} shows the result of this Clustering algorithm. 
We start with a sub-path which is similar in structure to the sub-path between the stop signs in Fig. \ref{fig:subpathdivision}. This sub-path is broken up into five segments as shown in Fig. \ref{fig:1Dto2DandSegmentationa}, three shown in blue corresponding to straight line part of the sub-path and two shown in red corresponding to turns in the sub-path. As shown in Fig \ref{fig:1Dto2DandSegmentationb} we use the speed limit of 11.1 m/sec for the straight line segments and 3.0 m/sec for the turn segments to determine the segment boundaries.

@@

% need better notation here.
In the consolidation step, a single set of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ 
%initial/final speeds, speed limit, peak acceleration and jerk 
must be calculated for each segment using the values corresponding to the waypoints comprising the segment.
The minimum of scalar across all constituent points is chosen.
While not time optimal, it does ensure that no constraints are violated while still keeping the number of total segments low.
In the present example, this means that in the straight portions of the subpath $v_m = SL$ and in the curves $v_m$ is set to the value required to obtain $a_{B,y}^{max}$ for the entirety of the curve.
As a result, many of the curves will have a constant speed.
For the sake of simplicity, the values of $a^+_{m,k}$, $a^-_{m,k}$, $j^+_{m,k}$, and $j^-_{m,k}$ are initially all set constant for the entire path, using values from \cite{Maurya2012,Hoberock1977,Long2000}.

A jerk vs. time profile is fit to each segment using the algorithm described in the next section.
Speed and acceleration must be continuous between each segment, but only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
The values of $a_i$ and $v_i$ for the first segment are set to the value currently reported from sensor data.
Since the final acceleration is to be set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments. 
Primary goal of this process is the continuity of position, speed, and acceleration.

@@

To introduce the process of computing a jerk profile for each segment, let us consider the case in which a vehicle needs to travel some arbitrary distance ahead.
The vehicle is moving at an initial speed and needs to accelerate to a traveling speed, maintain that speed for some distance, then brake to a stop at a prescribed end location.
Figure \ref{fig:full7phasespec} shows this trajectory over time.

\begin{figure}[tb]
  \centering
  \includegraphics[width=1.0\columnwidth]{graphics/Full7PhaseSpecVertical.png}
  \caption{
    7-phase constant jerk profile along curvilinear distance axis.
    All values needed to completely parameterize a single path segment for trajectory are labeled here.}
  \label{fig:full7phasespec}
\end{figure}

@@

We propose to define 4 other basic profiles to cover all feasible values in $\mathbf{b}$, and our algorithm judiciously chooses the appropriate profile. These 4 profiles cover cases in which the 7-phase solution fails. For instance, if $L$ is too short or $v_m$ is too high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
%The other profiles are explained below.
All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, traveling speed of zero, negative speeds in $\mathbf{b}$, etc).
It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
To find closed-form solutions for $[\Delta t_1, ..., \Delta t_M]$, given $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$, a symbolic solver is used 
%\textbf{RG: can specify solver name???} .
These 4 profiles are:

@@

% \subsection{Stitching Segment Solutions Together} \label{sec:unifiedsolution}
For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
This results in a negative time interval for one or more phases.
In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
Specifically, jerk may be increased while peak acceleration is decreased.
This process is referred to herein as ``accel-jerk tuning''.
If no acceptable values are found within the valid ranges of $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$, a new jerk profile must be chosen.

@@

For each segment with $M$ jerk phases, the following values are computed, which correspond to the instants at which jerk changes:
\begin{itemize}
  \item $M$ time durations $\Delta t_{ref}$. These must be added to the last time value from the previous segment to translate them into reference time.
  \item $M$ jerk values $j(t_{ref})$, which may be appended directly to the preceding jerk values.
\end{itemize}
It is now possible to integrate $j(t)$ to determine an instantaneous trajectory at any moment in time.
